### 2021-04-19 papers/final/paper.tex
start individual progress report

%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf,authorversion]{acmart}
%% NOTE that a single column version may be required for
%% submission and peer review. This can be done by changing
%% the \doucmentclass[...]{acmart} in this template to
%% \documentclass[manuscript,screen,review]{acmart}

\begin{document}
\settopmatter{printacmref=false}
\setcopyright{none}
\renewcommand\footnotetextcopyrightpermission[1]{}
\pagestyle{plain}
\title{Musical Deep Learning}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Nhut Minh Phan}
\email{phan92@uw.edu}
\affiliation{%
  \institution{University of Washington}
  \city{Bothell}
  \state{WA}
  \country{USA}
}

\author{Alex Kyllo}
\email{akyllo@uw.edu}
\affiliation{%
  \institution{University of Washington}
  \city{Bothell}
  \state{WA}
  \country{USA}
}

\begin{abstract}
  The abstract will go here.
\end{abstract}

\keywords{deep learning, neural networks, music}

\maketitle

\section{Introduction}

Example citation \cite{huang_music_2018}.

\section{Methods}

\section{Results}

\section{Discussion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{paper}

\end{document}
\endinput