### 2021-04-24 papers/progress/jack/progress-jack.tex
Add jacks report

%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf,authorversion]{acmart}
%% NOTE that a single column version may be required for
%% submission and peer review. This can be done by changing
%% the \doucmentclass[...]{acmart} in this template to
%% \documentclass[manuscript,screen,review]{acmart}

\begin{document}
\settopmatter{printacmref=false}
\setcopyright{none}
\renewcommand\footnotetextcopyrightpermission[1]{}
\pagestyle{plain}
\title{Progress Report for CSS 586: \\
Music Transcription}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

\author{Jack Phan}
\email{phan92@uw.edu}
\affiliation{%
  \institution{University of Washington}
  \city{Bothell}
  \state{WA}
  \country{USA}
}

\begin{abstract}

\end{abstract}

\keywords{deep learning, neural networks, music}

\maketitle

\section{Introduction}

\section{Related Work}

\section{Planned Methods}

\section{Progress}

\section{Future Work}

\bibliographystyle{ACM-Reference-Format}
\bibliography{progress-jack}

\end{document}
\endinput