### 2021-05-01 papers/progress/alex/progress-alex.tex
progress paper work

%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf,authorversion]{acmart}
%% NOTE that a single column version may be required for
%% submission and peer review. This can be done by changing
%% the \doucmentclass[...]{acmart} in this template to
%% \documentclass[manuscript,screen,review]{acmart}

\begin{document}
\settopmatter{printacmref=false}
\setcopyright{none}
\renewcommand\footnotetextcopyrightpermission[1]{}
\pagestyle{plain}
\title{Progress Report for CSS 586 Course Project: \\
Modeling Latent Patterns in Music}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

\author{Alex Kyllo}
\email{akyllo@uw.edu}
\affiliation{%
  \institution{University of Washington}
  \city{Bothell}
  \state{WA}
  \country{USA}
}

\begin{abstract}
This report explores recent research in symbolic music modeling with
deep learning and provides a progress report on a course project to
train a generative model of classical and popular music on the
MusicNet and Lakh MIDI datasets.
\end{abstract}

\keywords{deep learning, sequence learning, generative modeling,
  recurrent neural networks, music, MIDI}

\maketitle

\section{Introduction}

Machine learning models of music have interesting applications in
music information retrieval and creative tools for musical artists and
educators. Generative models can create accompaniments for music,
transfer styles between clips of music, and even generate entirely new
music. Music is challenging to model because it exhibits a complex
hierarchy of recurring patterns and long-range temporal dependencies,
and because both musical scores and performances have multiple
possible digital representations.

Depending on the task, machine learning models of music may be trained
on the audio signal itself, either in a time domain or a frequency
domain representation, or they may be trained on a digital symbolic
representation of music, the most common of which is MIDI (Musical
Instrument Digital Interface) notation. MIDI is an encoding of music
as streams of bytes in one or more tracks or channels, each
representing a sequence of 128 possible pitch values, along with
timing, pressure and instrument values. A music transcription model
may transcribe an audio signal as a MIDI score, which can easily be
converted into other symbolic representations such as sheet music for
human performers to read from, while a synthesizer model can convert
MIDI representations into audio signals. A generative music model can
be either trained to generate raw audio, or trained to produce a score
that must be played by either a synthesizer or by humans to produce an
audio music performance.

\section{Related Work}

The state of the art in music generation has a long way to go before
it can consistently generate music scores or performances that would
be enjoyable and popular to listen to. There are a number of
interesting, recent papers and projects in the area of music machine
learning, several of which focus on learning long-term dependencies in
music to produce longer clips with coherent structure.

Google's \href{https://magenta.tensorflow.org/}{Magenta} is an
umbrella project for music deep learning research and development of
software tools to expose these models for use by creative artists and
students.

MusicVAE, part of the Magenta project, is a variational LSTM
autoencoder for MIDI that incorporates a novel hierarchical structure
using a ``conductor'' recurrent layer in its decoder model to better
capture structure at multiple levels and avoid the problem of
``posterior/mode collapse'' whereby a generative model learns to
ignore its latent code and rely on autoregression
\cite{roberts_hierarchical_2018}.

Another Magenta model called Music Transformer is a generative model
that borrows its approach from the Natural Language Processing (NLP)
domain, using a self-attention network to model MIDI music as a
sequence of discrete tokens with relative positional dependencies
\cite{huang_music_2018}. This model was trained on the MAESTRO dataset
of Piano-e-competition performances \cite{hawthorne2019enabling} and
its generated piano music received favorable qualitative (Likert
scale) ratings from human listeners \cite{huang_music_2018}.

MuseGAN \cite{dong2017musegan} is an application of Generative Adversarial
Networks to polyphonic MIDI music generation, trained on four-bar phrases of a
multi-track pianoroll representation of rock songs from the Lakh Midi Dataset
\cite{raffel_learning-based_2016}.

A major advantage of working with the symbolic representation of music is that
it is of far lower dimensionality than the raw audio waveforms of a recorded
performance, which makes it less computationally expensive. However, there are
many aspects of musical performance that are not captured by a symbolic
representation, so the expressiveness of symbolic generative models is
constrained \cite{manzelli_conditioning_2018}.

Other research has focused on modeling raw audio waveforms directly. WaveNet is
a causal convolutional neural network for generating raw audio waveforms,
developed by Google DeepMind, which achieves state of the art performance in
generating natural sounding speech from text, but is also capable of generating
short, realistic snippets of audio music \cite{oord_wavenet_2016}.

Another model named SampleRNN generates raw audio waveforms using a three-tier
hierarchy of gated recurrent units (GRU) to model recurrent structure at
multiple temporal resolutions \cite{mehri_samplernn_2017}.

Jukebox by OpenAI utilizes a vector-quantized variational autoencoder
(VQ-VAE) to compress raw audio into a sequence of discrete codes and
models these sequences using autoregressive transformers to generate
music of various popular genres, including singing and lyrics
\cite{dhariwal2020jukebox}.

Audio data can also be modeled in the frequency domain through the use of
Fourier analysis. The recent MelNet model is trained on spectrograms and can
learn musical structures such as melody and harmony and variations in volume,
timbre and rhythm \cite{vasquez2019melnet}.

Prior work points out that the division between symbolic music notes
and the sounds of music is analogous to the division between symbolic
language and utterances in speech, which may inspire ideas for
combining the two approaches \cite{hawthorne2019enabling}. A paper
from Boston University describes an effort to combine the symbolic and
waveform approaches to music modeling, by training an LSTM to learn
melodic structure of different styles of music, then providing
generations from this model as conditioning inputs to a WaveNet-based
raw audio generator \cite{manzelli_conditioning_2018}.

\section{Planned Methods}

This research project will focus on the generative modeling of
symbolic music using MIDI data as inputs, because of the advantages of
symbolic music models in representing long-term structure in musical
compositions to produce generations with coherent structure and use of
repetition over long time scales.

\subsection{Datasets}

Two primary datasets will be used for this research project: MusicNet,
which is a collection of 330 freely licensed European classical music
recordings with aligned MIDI scores \cite{thickstun2017learning}, and
the Lakh MIDI Dataset, which includes a collection of 45,129 Creative
Commons licensed MIDI files of popular music songs that have been
matched and aligned to MP3 recordings, and of which we will use a
subset for model training and evaluation
\cite{raffel_learning-based_2016}.

\subsection{Data Preprocessing}

Several choices must be made in how to preprocess binary MIDI files
into training examples for a neural network. There are multiple
open-source Python packages capable of reading MIDI files into NumPy
array-based representations, such as pretty\_midi
\cite{raffel_pretty_midi_2014} and Pypianoroll
\cite{dong_pypianoroll_2018}. In order to accommodate polyphonic
music, we will convert each MIDI file into a pianoroll representation,
wherein each instrument track is a sparse matrix that multi-hot
encodes the velocity values for each of 128 possible pitch levels at
each timestep. Because songs are typically at least a few minutes long
and of varying length, it will not be feasible to train with entire
songs as examples, so we will crop songs into phrases of equal numbers
of measures, to use as training data.

The result of this preprocessing is that each training example will be
a 3D tensor of shape (tracks x notes x ticks) and stacking the
training examples will produce a 4D tensor.

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{first_four_bars.png}
  \caption{Pianoroll visualization of the first four measures of
    Beethoven's Serioso String Quartet}
  \Description{A pianoroll representation of cello, viola and violin
    parts from a Beethoven piece.}
\end{figure}

Data augmentation is also possible and we plan to test its impact on
our results--the literature suggests augmentation via pitch shifting
the entire training example up or down by up to six semitones, and
increasing or reducing the speed by up to 10\% in order to create
additional training examples and reduce overfitting
\cite{oore_this_2018}.

\subsection{Model Fitting}

We will explore several modeling approaches to generating symbolic music:

\begin{itemize}
  \item{Sliding window sequence prediction with RNNs (LSTM/GRU)}
  \item{Sliding window sequence prediction with Transformers}
  \item{Latent space interpolation with Sequential VAEs}
  \item{Latent space interpolation with Sequential GANs}
\end{itemize}

\subsection{Model Evaluation}

Evaluation of generative models is challenging because there is no
equivalent of an accuracy metric like what is used in supervised
learning. For autoencoder models we can measure how accurately the
model can reconstruct its own inputs, but this does not tell us the
quality of the interpolated examples. Generative models are typically
evaluated using a combination of qualitative metrics whereby human
judges rate the quality of the generated examples (essentially a
Turing test), and quantitative metrics that assess the differences in
the parametric distributions of generated and real examples. Yang and
Lerch (2020) proposes a set of metrics informed by music theory, for
probabilistically evaluating how similar the generations are to known
sample distributions of real music \cite{yang_evaluation_2020}. These
metrics include counts, ranges, histograms and transition matrices of
pitches and note lengths, then the Kullback-Leibler divergence and
overlapping area of the probability density functions are use to
compare against known reference distributions per musical genre
\cite{yang_evaluation_2020}. Due to the cost and time requirements
associated with designing a human subjects experiment, we plan to
utilize this quantitative approach to generation quality assessment.

\section{Progress and Remaining Work}

At this stage, we have decided on the modeling problem, downloaded the
MIDI datasets, and begun exploring them to learn how to interpret,
process, and visualize them. We have started writing a training data
processing pipeline that will provide batches of training examples to
Keras model layers for fitting.

\bibliographystyle{ACM-Reference-Format}
\bibliography{progress-alex}

\end{document}
\endinput