### 2021-06-08 papers/final/alex/paper.tex
regenerate notebooks from updated scripts

%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf,authorversion]{acmart}
\usepackage{pgf}
\usepackage{pgfplots}
\usepackage{url}
%% NOTE that a single column version may be required for
%% submission and peer review. This can be done by changing
%% the \doucmentclass[...]{acmart} in this template to
%% \documentclass[manuscript,screen,review]{acmart}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}
\settopmatter{printacmref=false}
\setcopyright{none}
\renewcommand\footnotetextcopyrightpermission[1]{}
\pagestyle{plain}
\title{Modeling the Latent Space of Symbolic String Quartet Music}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

\author{Alex Kyllo}
\email{akyllo@uw.edu}
\affiliation{%
  \institution{University of Washington}
  \city{Bothell}
  \state{WA}
  \country{USA}
}

\begin{abstract}
  This paper presents the results of a project to apply a recurrent
  variational autoencoder to the task of modeling a latent space for
  generating pieces of multi-instrument symbolic classical music
  through linear interpolation. The code utilized for this research is
  available at: \url{https://github.com/jackPhn/CSS586_Project/}
\end{abstract}

\keywords{deep learning, recurrent neural networks, sequential models,
  autoencoders, music modeling, latent space modeling, generative
  modeling}

\maketitle

\section{Introduction}

Machine learning models of music have interesting applications in
music information retrieval and creative tools for musical artists and
educators. Generative models can create accompaniments for music,
blend and transfer styles between clips of music, and even generate
entirely new music. Music is challenging to model because it is
inherently high-dimensional, exhibiting a complex hierarchy of
recurring patterns and long-range temporal dependencies, and because
musical scores have multiple possible digital representations with
distinct advantages and disadvantages.

Depending on the task, machine learning models of music may be trained
on the audio signal of a musical performance, either in a time domain
or a frequency domain representation, or they may be trained on a
digital symbolic representation of music, the most common of which is
MIDI (Musical Instrument Digital Interface) notation. MIDI is an
encoding of music as streams of bytes in one or more tracks or
channels, each representing a sequence of 128 possible pitch values
(where 0 is the lowest pitch and 127 is the highest), along with
timing, pressure and instrument identifier values. A generative
symbolic music model can produce a symbolic score in MIDI format,
which must be played by synthesizer software or by humans to produce
an audio music performance. This project focuses on generative
modeling of symbolic (MIDI) music to compose original musical scores
by blending input scores through continuous latent space interpolation.

A survey of related works in generative modeling of polyphonic music
turned up several models focused on multi-instrument pop/rock music
arrangements or classical piano performances. In contrast, this
project focuses on variational autoencoding and latent space
interpolation of classical string quartets (arrangements for an
ensemble of two violins, one viola and one cello) and is, to the
author's knowledge, the first project to apply this modeling approach
to this type of musical arrangement.

\section{Related Work}

The state of the art in music generation still has a long way to go
before it can consistently generate music scores or performances that
would be broadly enjoyable for humans to listen to, but for this
reason it is an area of significant opportunity, where a number of
recent research projects have shown promising progress.

Google's \href{https://magenta.tensorflow.org/}{Magenta} is an
umbrella project for music deep learning research and development of
software tools to expose these models for use by creative artists and
students. MusicVAE, part of the Magenta project, is a variational Long
Short-Term Memory (LSTM) autoencoder for MIDI that incorporates a
novel hierarchical structure using a ``conductor'' recurrent layer in
its decoder model to better capture structure at multiple levels and
avoid the problem of ``posterior/mode collapse'' whereby a generative
model learns to ignore its latent codes and rely on autoregression to
generate output sequences \cite{roberts_hierarchical_2018}. This model
is trained on 16-bar paragraphs of music and is capable of generating
new melodies that blend two given melodies together via latent space
interpolation. Among the literature surveyed, MusicVAE's approach is
the most similar to this work; in comparison MusicVAE dedicates
significant effort to devising an architecture that allows the model
to generate very long sequences of notes, which is beneficial for
other generation tasks but not strictly necessary for
interpolation. Our model is significantly simpler and demonstrates
that interpolation works well even on a single bar of music at a time.

Another Magenta model called Music Transformer is a generative model
that borrows its approach from the Natural Language Processing (NLP)
domain, using a self-attention network to model MIDI music as a
sequence of discrete tokens with relative positional dependencies
\cite{huang_music_2018}. The focus of this model is on learning
long-term dependencies in music to produce longer clips of music with
coherent structure. Music Transformer was trained on a dataset of
Piano-e-competition performances \cite{hawthorne2019enabling} and its
generated piano music received favorable qualitative (Likert scale)
ratings from human listeners for its resemblance to human-composed
music \cite{huang_music_2018}. In contrast to MusicVAE and this work,
Music Transformer is capable of generating much longer sequences of
music, but does so via sequence extrapolation rather than
interpolation, so it can continue a given priming melody but does not
blend multiple given melodies together. It also assumes a single
instrument track (piano).

MuseGAN \cite{dong2017musegan} is an application of Generative
Adversarial Networks (GAN) to polyphonic MIDI music generation,
trained on four-bar phrases of a multi-track pianoroll representation
of rock songs from the Lakh Midi Dataset
\cite{raffel_learning-based_2016}. Like MusicVAE, MuseGAN includes a
two-level generator that first samples latent codes at the phrase or
bar level, then generates notes within the bars, to produce
longer-term structural patterns.

A major advantage of working with the symbolic representation of music
is that it is of far lower dimensionality than the raw audio waveforms
of a recorded performance, which makes it less computationally
expensive. However, there are many stylistic aspects of musical
performance that are not captured by a symbolic representation, and
may be specific to a particular performer, so the expressiveness of
symbolic generative models is limited in comparison
\cite{manzelli_conditioning_2018}.

Other research has focused on modeling raw audio waveforms directly. WaveNet is
a causal convolutional neural network for generating raw audio waveforms,
developed by Google DeepMind, which achieves state of the art performance in
generating natural sounding speech from text, but is also capable of generating
short, realistic snippets of audio music \cite{oord_wavenet_2016}.
Another model named SampleRNN generates raw audio waveforms using a three-tier
hierarchy of gated recurrent units (GRU) to model recurrent structure at
multiple temporal resolutions \cite{mehri_samplernn_2017}.

Prior work points out that the division between symbolic music notes
and music performances, is analogous to the division between symbolic
language and utterances in speech, which may inspire ideas for
combining the two approaches \cite{hawthorne2019enabling}. A paper
from Boston University describes an effort to combine the symbolic and
waveform approaches to music modeling, by training an LSTM to learn
melodic structure of different styles of music, then providing
generations from this model as conditioning inputs to a WaveNet-based
raw audio generator \cite{manzelli_conditioning_2018}.

While there may be significant opportunity in future approaches that
combine the two, contemporary research generally treats symbolic and
audio music modeling as separate problem domains and this work does
the same, choosing to focus on symbolic music.

\section{Methods}

\subsection{Datasets}

The dataset used for this research project is MusicNet, which is a
collection of 330 freely licensed European classical music recordings
with aligned MIDI scores \cite{thickstun2017learning}.  The model is
trained on 36 string quartets (four-part arrangements with two
violins, one viola and one cello) by composers Bach, Beethoven,
Dvorak, Haydn, Mozart, and Ravel.

\subsection{Data Preprocessing}

Several choices must be made in how to preprocess binary MIDI files
into training examples for a neural network. There are multiple
open-source Python packages that assist with the process of reading
MIDI files from their binary on-disk representations into Python
objects, such as pretty\_midi \cite{raffel_pretty_midi_2014},
Pypianoroll \cite{dong_pypianoroll_2018} and music21
\cite{cuthbert_music21_2010}.

In order to accommodate polyphonic music, each MIDI file is converted
into a modified pianoroll representation, an example of which is
visualized in Figure \ref{pianoroll}. While the standard pianoroll
represents each time-step as a multi-hot encoded vector of note
values, because most notes are not being played at most timesteps
there is an extreme class imbalance and sparsity in the matrix. This
is reported in Kumar and Ravindran (2019) \cite{kumar2019polyphonic}
to cause problems in model fitting and metrics calculation, so they
introduce an alternative ``multi-stream'' representation of the
pianoroll, which we closely, (but not exactly) emulate in this work.

In our modified representation, each instrument track is a sequence of
integers from 0-129 representing the MIDI note pitch value being
played at the given time step, where values 0-127 are pitches, 128 is
a rest (silence) and 129 is sustain, indicating that the previously
played pitch is held continuously. MIDI data contains pressure values
that indicate how hard or soft each note is played; this information
is discarded in our preprocessing and all notes are represented with
constant pressure. Because chords (multiple notes being played
simultaneously by the same instrument) are relatively rare in string
quartets, the lowest note of each chord is taken and the rest
discarded, to reduce dimensionality. Because notes at the extreme low
and high end are rare and are out of range of common instruments, the
note pitch values are clipped to a narrower range and ordinal
encoded. For the 36 string quartets found in the MusicNet database,
this results in 66 distinct values, where 0-63 represent notes played
by the violins, viola and cello, 64 is rest and 65 is sustain. Figure
\ref{pitch_dist} depicts the distribution of MIDI pitch values per
instrument and the degree to which they overlap.

Our software package also includes functions to convert this
representation back into a Music21 Score object, which can then be
written to a MIDI file that can be synthesized for human listening.

\begin{figure}[htbp]
    \begin{center}
        \scalebox{0.6}{\input{figures/pitch_dist.pgf}}
    \end{center}
    \caption{Distribution of pitch values per instrument in the MusicNet string quartets}
    \label{pitch_dist}
\end{figure}

Because songs are typically at least a few minutes long and of varying
length, it is not feasible to train with entire songs as examples, so
we crop songs into phrases of equal numbers of beats to use as
training data. The result of this preprocessing is that each training
example is a 2D matrix of shape (time steps x instrument tracks) and
stacking the training examples produces a 3D tensor. Information
regarding which track corresponds to which musical instrument is lost
and therefore must be maintained separately to reconstruct the MIDI
representation. For string quartets this instrument ensemble is always
\texttt{[40, 40, 41, 42]}, which is an array of MIDI instrument codes
representing two violins, one viola and one cello. Tempo information
is also lost from this representation, which uses symbolic time steps
rather than wall clock time, so the outputs are standardized to the
default tempo of 120 beats per minute.

As the model will produce one output per instrument track, it can
incorporate only a fixed selection of instrument parts, similar to how
MusicVAE models three-part (drum, bass and melody)
\cite{roberts_hierarchical_2018} and MuseGAN models five-part (drum,
bass, guitar, string, piano) arrangements. Therefore the number of
tracks is a hyperparameter to the network architecture definition and
this model is trained for string quartets, which are the most common
type of arrangement found in the MusicNet dataset.

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{figures/first_bar.png}
  \caption{Pianoroll visualization of the first measure of
    Beethoven's Serioso String Quartet in F Minor}
  \label{pianoroll}
  \Description{A pianoroll representation of cello, viola and violin
    parts from a Beethoven piece.}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{figures/first_bar_sheet.png}
  \caption{Sheet music of the first measure of Beethoven's Serioso
    String Quartet in F Minor}
  \label{sheet}
  \Description{A measure of sheet music cello, viola, and violin parts
    from Beethoven piece.}
\end{figure}

\subsection{Model Design}

The model is a recurrent variational autoencoder; both Long Short-Term
Memory (LSTM) and Gated Recurrent Unit (GRU) cell types were tested
but this choice did not make a significant difference in the results.

The encoder model utilizes an embedding layer to encode the note pitch
integers as vectors; embedding vector length is a tunable
hyperparameter and we found the best results with it set to 8. Then
two LSTM layers, with a dropout layer in between, convert the
sequences into latent codes, which are modeled as a 512-dimensional
Gaussian distribution $z$, produced by the encoder output Lambda layer
(Figure \ref{decoder}).

The decoder model samples from the latent space $z$ with the
distribution parameterized by the $\mu$ and $log(\sigma)$ values
learned by the encoder, and then utilizes two LSTM layers with a
dropout layer between them to generate sequences from the latent
codes. Then the network splits into four parallel outputs, one per
instrument track, where a fully connected layer with softmax
activation is used to select the most probable note value for the
the current track at each time step (Figure \ref{decoder}).

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{figures/encoder.png}
  \caption{A directed graph diagram of the encoder network.}
  \label{encoder}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{figures/decoder.png}
  \caption{A directed graph diagram of the decoder network (only one
    of four parallel output layers shown).}
  \label{decoder}
\end{figure}

The autoencoder reconstruction loss function is sparse categorical
cross-entropy, which is added to the Kullback-Leibler divergence of
the approximate posterior distribution from a multivariate Gaussian
distribution with mean 0 and standard deviation 1, resulting in the
(negative) evidence lower bound (ELBO) loss (Equation \ref{nelbo}),
which is a standard variational autoencoder loss function. This loss
function is applied to each track independently and then summed across
all four tracks with equal weight.

\begin{equation}
  \label{nelbo}
\mathbb{E}[log{p_\theta}(x|z)]-\mathbb{KL}(q_\lambda(z|x)||p(z))
\end{equation}

where $\lambda$ and $\theta$ represent the encoder and decoder
parameters respectively, while $q_\lambda(z|x)$ represents the
encoder, which approximates the true posterior probability $p(z|x)$,
and $p_\theta(x|z)$ represents the decoder, which parameterizes the
likelihood $p(x|z)$ \cite{roberts_hierarchical_2018}.

\subsection{Model Experimentation}

The following hyperparameters were tuned, with tested values listed
and best parameters in \textbf{bold}:

\begin{itemize}
  \tightlist
  \item batch size: \textbf{32}, 64, 96
  \item learning rate: 0.001, \textbf{0.0002}
  \item recurrent layer type: LSTM, GRU (made no difference)
  \item recurrent encoder layer directions: \textbf{unidirectional}, bidirectional
  \item recurrent layer cells: 64, \textbf{128}, 256
  \item latent code dimension: 128, 256, \textbf{512}
  \item embedding dimension: 4, \textbf{8}, 16
  \item dropout rate: 0.2, 0.4, \textbf{0.5}
  \item time steps per beat (quarter note): \textbf{4}, 8
  \item beats per phrase: \textbf{4}, 8
\end{itemize}

Each model was trained for up to 500 epochs with an early stopping
patience of 20 epochs. The model was cross-validated on 10\% of the
training data after each epoch. The best model was selected by a
combination of validation loss and listening to hand-picked sample
reconstructions and interpolations.

Data augmentation was also tested--Oore et al (2018)

\cite{oore_this_2018} suggests augmentation via pitch shifting each
training example up or down by up to six semitones in order to create
additional training examples and reduce overfitting, so we followed
that recommendation and used it to double the size of the training
dataset. As a result, we saw much lower validation losses (compare
Figure \ref{loss} and Figure \ref{loss_augment}) and a small but
noticeable improvement in the fidelity of the song reconstructions,
suggesting that this method was effective in reducing overfitting.

\begin{figure}[htbp]
    \begin{center}
        \scalebox{0.5}{\input{figures/loss.pgf}}
    \end{center}
    \caption{Training and validation loss for the best performing model - without data augmentation}
    \label{loss}
\end{figure}

\begin{figure}[htbp]
    \begin{center}
        \scalebox{0.5}{\input{figures/loss_augment.pgf}}
    \end{center}
    \caption{Training and validation loss for the best performing model - with data augmentation}
    \label{loss_augment}
\end{figure}

\section{Results}

The reconstructed music produced by the model sounds reminiscent of
the original inputs, preserving the key signature and general rhythmic
and melodic structure, albeit with small but noticeable differences at
the individual note level, such as more silence and less variation in
pitch.

A collection of samples of the autoencoded outputs and interpolations,
converted to mp3 format, is presented at the following URL:
\url{https://drive.google.com/drive/folders/1SlQOI_cL452PftO3KmOYO_qglE2RotCX?usp=sharing}.
Each of the 18 directories contains five interpolations between two of
the string quartet pieces in the MusicNet corpus, numbered 0-4, where
0 is the reconstruction of the first piece, 4 is the reconstruction of
the second piece, and 1-3 are linear interpolations between the two
pieces. Each file is cropped to the first 16 measures only.

The interpolations are between the following 18 pairs of pieces
denoted by their filenames in the original MusicNet reference MIDI
archive, which can be downloaded from:
\url{https://homes.cs.washington.edu/~thickstn/media/musicnet_midis.tar.gz}

\begin{itemize}
\tightlist
\item \texttt{Haydn/2104\_op64n5\_1.mid} \& \texttt{Ravel/2178\_gr\_rqtf2.mid}
\item \texttt{Haydn/2105\_op64n5\_2.mid} \&   \texttt{Ravel/2179\_gr\_rqtf3.mid}
\item \texttt{Haydn/2106\_op64n5\_3.mid} \&   \texttt{Ravel/2177\_gr\_rqtf1.mid}
\item \texttt{Beethoven/2497\_qt11\_4.mid} \& \texttt{Mozart/1822\_kv\_421\_1.mid}
\item \texttt{Beethoven/2433\_qt16\_3.mid} \& \texttt{Mozart/1859\_kv\_464\_2.mid}
\item \texttt{Beethoven/2368\_qt12\_4.mid} \& \texttt{Mozart/1807\_kv\_387\_3.mid}
\item \texttt{Beethoven/2314\_qt15\_2.mid} \& \texttt{Mozart/1791\_kv\_465\_4.mid}
\item \texttt{Beethoven/2480\_qt05\_1.mid} \& \texttt{Mozart/1792\_kv\_465\_1.mid}
\item \texttt{Beethoven/2481\_qt05\_2.mid} \& \texttt{Mozart/1835\_kv\_590\_3.mid}
\item \texttt{Beethoven/2379\_qt08\_4.mid} \& \texttt{Mozart/1805\_kv\_387\_1.mid}
\item \texttt{Beethoven/2365\_qt12\_1.mid} \& \texttt{Mozart/1793\_kv\_465\_2.mid}
\item \texttt{Beethoven/2562\_qt02\_4.mid} \& \texttt{Mozart/1790\_kv\_465\_3.mid}
\item \texttt{Beethoven/2494\_qt11\_1.mid} \& \texttt{Mozart/1789\_kv\_465\_2.mid}
\item \texttt{Beethoven/2403\_qt01\_4.mid} \& \texttt{Mozart/1788\_kv\_465\_1.mid}
\item \texttt{Beethoven/2376\_qt08\_1.mid} \& \texttt{Dvorak/1916\_dvq10m1.mid}
\item \texttt{Beethoven/2384\_qt13\_4.mid} \& \texttt{Bach/2242\_vs1\_2.mid}
\item \texttt{Beethoven/2560\_qt02\_2.mid} \& \texttt{Beethoven/2621\_qt07\_1.mid}
\item \texttt{Beethoven/2377\_qt08\_2.mid} \& \texttt{Beethoven/2381\_qt13\_1.mid}
\end{itemize}

\subsection{Generated Sample Evaluation}

Evaluation of generative models is challenging because there is no
direct equivalent of an accuracy or error metric like what is used in
supervised learning. For autoencoder models we can measure how
accurately the model can reconstruct its own inputs, but this does not
tell us the quality of the interpolated examples. Generative models
are typically evaluated using a combination of qualitative metrics
whereby human judges rate the quality of the generated examples and
guess whether they were generated by a machine (essentially a Turing
test), and quantitative metrics that assess the differences in the
parametric distributions of generated and real examples. Yang and
Lerch (2020) propose a set of metrics informed by music theory, for
probabilistically evaluating how similar the generations are to known
sample distributions of real music \cite{yang_evaluation_2020}. These
metrics include counts, ranges, histograms and transition matrices of
pitches and note lengths, then compare the probability density or mass
functions of generated samples against that of known genuine source
samples \cite{yang_evaluation_2020}. Due to the cost and time
requirements associated with designing a human subjects experiment for
qualitative evaluation, we utilize this quantitative approach to the
assessment of the pitch distributions of generated samples.

Figure \ref{pitch_counts} depicts histograms of source and generated
music samples respectively (1 sample = 1 measure) by the number of
distinct pitch counts they contain. In both samples the mode is 9
distinct pitches, but the source arrangements have a median of 11
while the generations have a median of 8. This supports our own
qualitative observation that the generated samples have slightly less
pitch diversity than the genuine compositions.

\begin{figure}[htbp]
    \begin{center}
        \scalebox{0.6}{\input{figures/pitch_count_histogram.pgf}}
    \end{center}
    \caption{Distinct pitch counts per sample for generated and source samples.}
    \label{pitch_counts}
\end{figure}

Figure \ref{pch} depicts the source and generated samples' proportion
of notes by each of the 12 notes in the chromatic scale, regardless of
octave, where 0 = C, 1 = C$\sharp$/D$\flat$, 2 = D, and so on. These
distributions are very similar, with the generated samples having more
Cs and fewer Ds than the source samples.

\begin{figure}[htbp]
    \begin{center}
        \scalebox{0.6}{\input{figures/pitch_class_histogram.pgf}}
    \end{center}
    \caption{Comparative pitch class histogram for source and generated samples.}
    \label{pch}
\end{figure}

Figures \ref{source_ranges} and \ref{generated_ranges} depict the
distribution of pitch range (highest pitch - lowest pitch) among
samples. The distributions appear similar; the source distribution has
a mean of 30.5 while the generated distribution has a mean of 26.3,
(the medians are similar to the means for both) so the generated
distributions have a slightly narrower pitch range.

\begin{figure}[htbp]
    \begin{center}
        \scalebox{0.6}{\input{figures/pitch_range_source.pgf}}
    \end{center}
    \caption{Pitch range per sample for source samples.}
    \label{source_ranges}
\end{figure}

\begin{figure}[htbp]
    \begin{center}
        \scalebox{0.6}{\input{figures/pitch_range_generated.pgf}}
    \end{center}
    \caption{Pitch range per sample for generated samples.}
    \label{generated_ranges}
\end{figure}

Figures \ref{source_interval} and \ref{generated_interval} describe
the distributions average pitch intervals between consecutively played
notes; these distributions are very similar, both are centered with
approximate mean of 0 and standard deviation of 1.5.

\begin{figure}[htbp]
    \begin{center}
        \scalebox{0.6}{\input{figures/avg_pitch_interval_source.pgf}}
    \end{center}
    \caption{Average pitch interval per sample for source samples.}
    \label{source_interval}
\end{figure}

\begin{figure}[htbp]
    \begin{center}
        \scalebox{0.6}{\input{figures/avg_pitch_interval_generated.pgf}}
    \end{center}
    \caption{Average pitch interval per sample for generated samples.}
    \label{generated_interval}
\end{figure}

Finally, figures \ref{pctm_source} and \ref{pctm_generated} depict
heatmaps of transition probability between the twelve pitch classes in
the chromatic scale. The source samples exhibit higher probabilities
of transition between different notes, while in the generated sample
there is more repetition of the same note, again confirming the
relative lack of pitch diversity in the generated samples. However,
the dark areas of very low probability in the source distribution are
similarly dark in the generated distribution; the model is unlikely to
generate pitch transitions that are rare in the source distribution.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=\linewidth]{figures/pctm_source.png}
  \caption{Pitch class transition matrix for source samples.}
  \label{pctm_source}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=\linewidth]{figures/pctm_generated.png}
  \caption{Pitch class transition matrix for generated samples.}
  \label{pctm_generated}
\end{figure}

\section{Discussion}

Our research demonstrates that a relatively simple LSTM variational
autoencoder network can learn a continuous latent space for generating
realistic linear interpolations between symbolic, classical string
quartet pieces, which are a notably complex genre of music.
Furthermore, our model can achieve this while working
measure-by-measure, on sequences just 16 time steps in length, without
the need for a highly complex neural architecture capable of
memorizing very long sequences. Listening to the generated samples and
profiling their pitch distributions reveals that some of the pitch
diversity of the original music is lost, but the overall melodic and
rhythmic structure still generally resemble the originals. It was our
aim to produce interpolations that retained enough aesthetic quality
to be pleasing to listen to and useful in inspiring the creative
process, and we believe our work meets that goal.

Future work in this domain could involve disentanglement of the latent
dimensions to enable music style transfer and semantic editing via
interpolation in specific directions within the latent space. The VAE
model could also be adapted to other musical genres aside from string
quartets, to interpolate within or across genres.

\bibliographystyle{ACM-Reference-Format}
\bibliography{paper}

\end{document}
\endinput